
---
title: 'CENG 3516 Statistical Computing Final Report'
author:
- Cansel Hatice Küçükyılmaz - 170709049
- Muğla Sıtkı Koçman University Faculty of Engineering, Computer Engineering
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
  pdf_document:
    template: null

---
<h1> 1) Istanbul Metropolitan Municipality Student Aid </h1>
<h1>a. Data collection </h1>
<h2>Description of Data </h2>
This dataset contains information about the students who received educational assistance over the years. In this data source, it includes the number of students given educational aid by the Istanbul Metropolitan Municipality.

<html>
<head>
<style>
hr.new3 {
  border-top: 1px dashed red;
}

table {
  font-family: arial, sans-serif;
  border-collapse: collapse;
  width: 100%;
}

td, th {
  border: 1px solid #dddddd;
  text-align: left;
  padding: 8px;
}

tr:nth-child(even) {
  background-color: #dddddd;
}
</style>
</head>
<body>

<h2>Dataset Information</h2>

<table>
   <tr>
    <th>Column</th>
    <th>Description</th>
  </tr>
  <tr>
    <th>ASSISTANCE_YEAR</th>
    <th> It is the field that includes the year of education aid.</th>
  </tr>
  <tr>
    <td>ASSISTANCE_TERM</td>
    <td>It is the field that includes the period of education aid.</td>

  </tr>
  <tr>
    <td>UNIVERSITY</td>
    <td> Information on the school where students who receive educational aid study.</td>
  </tr>
  
  <tr>
    <td>AGE</td>
    <td>Ages of students receiving educational aid.</td>
  </tr>
  <tr>
    <td>COUNTY</td>
    <td> District knowledge of the students receiving educational aid.</td>
  </tr>
  <tr>
    <td>NUMBER_OF_STUDENTS</td>
    <td>Number of students receiving educational aid</td>
  </tr>

</table>
<b>Data Source:</b> https://data.ibb.gov.tr/dataset/8badf78e-2307-46cb-b3a1-83fc4e5d804a/resource/69484e6b-3440-4a61-bf59-19e5fafd49a9/download/dam_occupancy.csv
<br>
<b>License: Istanbul Metropolitan Municipality Open Data License </b>
<br>
<hr>

```{r}
library(ggplot2)
library(dplyr) 
```

```{r}
library(tidyverse)
library(sp) 
library(ggplot2)
library(ggplot2)
library(DescTools)

```

## Including Plots

<h1>b. Analyses </h1>

```{r}
tr1<- readRDS("gadm36_TUR_2_sp.rds") 
plot(tr1)
```
<br>
<br>
<br>
First of all, I would like to show the density of the students to whom IBB grants scholarships, by district, on the Istanbul district map.  For this, I first downloaded  turkey spatial data from GAMD. <https://gadm.org/download_country_v3.html>.  


```{r}
istanbul=tr1%>%subset(NAME_1=="Istanbul")
plot(istanbul)
```
<br>
<br>
<br>
In this turkey data, I pulled the istanbul data from the NAME_1 colum. And I had the map of Istanbul drawn by districts.

```{r}

istanbul@data$NAME_2 <- gsub("Arnavutkoy","Arnavutköy",istanbul@data$NAME_2 )
istanbul@data$NAME_2 <- gsub("Basaksehir","Başaksehir",istanbul@data$NAME_2 )
istanbul@data$NAME_2 <- gsub("Atasehir","Ataşehir",istanbul@data$NAME_2 )
istanbul@data$NAME_2 <- gsub("Basaksehir","Başakşehir",istanbul@data$NAME_2 )
istanbul@data$NAME_2 <- gsub("Çekmekoy","Çekmeköy",istanbul@data$NAME_2 )
istanbul@data$NAME_2 <- gsub("Beylikduzu","Beylikdüzü",istanbul@data$NAME_2 )

istanbul@data %>% as_tibble() %>% head(10) 
```
I saw that the district names were wrong in some of the data from the data set and I corrected them with the help of the gsub function.
```{r}
istanbul@data$NAME_2
```


```{r}
istdata<-fortify(istanbul)
head(istdata)
```

We need to convert the spatial data we have into a data.frame object, for this I made a dataframe with the fortify function.


```{r}
ibb_yardim <- read.csv("ibb-genc-universiteli.csv")
head(ibb_yardim)
```

I uploaded the Student Aid data that we pulled from IBB <https://data.ibb.gov.tr/dataset/8badf78e-2307-46cb-b3a1-83fc4e5d804a/resource/69484e6b-3440-4a61-bf59-19e5fafd49a9/download/dam_occupancy.csv> to the workspace with the help of the read.csv function. 


```{r}
ibb_yardim$COUNTY <- str_to_title(ibb_yardim$COUNTY)
head(ibb_yardim)

```

District names were entered in capital letters, I converted them to title format with the help of the str_to_title function.

```{r}
table(ibb_yardim$COUNTY)
```
Here, It seems that some county names are wrong.

```{r}
ibb_yardim$COUNTY<-gsub("Bağcilar" ,"Bağcılar",ibb_yardim$COUNT)
ibb_yardim$COUNTY<-gsub("Bakirköy" ,"Bakırköy",ibb_yardim$COUNT)
ibb_yardim$COUNTY<-gsub("Avcilar" ,"Avcılar",ibb_yardim$COUNT)
ibb_yardim$COUNTY<-gsub("Pendi̇k" ,"Pendik",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Sariyer" ,"Sarıyer",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Gazi̇osmanpaşa" ,"Gaziosmanpaşa",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Beyli̇kdüzü" ,"Beylikdüzü",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Bahçeli̇evler" ,"Bahçelievler",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Ümrani̇ye" ,"Ümraniye",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Fati̇h" ,"Fatih",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Zeyti̇nburnu" ,"Zeytinburnu",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Kağithane" ,"Kağıthane",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Sultanbeyli̇" ,"Sultanbeyli",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Cekmekoy" ,"Çekmeköy",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Sultangazi̇", "Sultangazi",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Şi̇şli̇", "Şişli",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Si̇li̇vri̇","Şişli",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Başakşehi̇r","Başakşehir",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Ataşehi̇r","Ataşehir",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Şi̇le","Şile",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Kadiköy","Kadıköy",ibb_yardim$COUNTY)
ibb_yardim$COUNTY<-gsub("Beşi̇ktaş","Beşiktaş",ibb_yardim$COUNTY)

```

I corrected the wrong county names with the help of the gsub function.


```{r}
year2019 <- ibb_yardim%>%filter(ASSISTANCE_YEAR==2019)
```

I generated a sub-dataframe for the students who applied in 2019.

```{r}
totalStudents_2019 <-  aggregate(NUMBER_OF_STUDENTS~COUNTY,year2019,sum)




head(totalStudents_2019%>%as_tibble)


```
The total number of scholarships earned by students by district.


```{r}
istanbul@data$NAME_2
```
County names from spatial data.

```{r}
students.number.based.on.county<- data_frame(id=rownames(istanbul@data), COUNTY=istanbul@data$NAME_2) %>% full_join(totalStudents_2019, by = "COUNTY")
```
Using full join, I combined the data we pulled from Spatial data and the data of students who received scholarships in 2019 from IBB based on county.


```{r}
head(students.number.based.on.county)
```




```{r}
final_map <- left_join(istdata, students.number.based.on.county, by = "id")
head(final_map)
```
join both data using left_join function.

```{r}
table(students.number.based.on.county$id)
```
```{r}
students.number.based.on.county

```




```{r}
str(students.number.based.on.county)
```


```{r}
students.number.based.on.county$id=as.character(as.numeric(students.number.based.on.county$id)+1)
```
There is a difference on ids so I added 1 to remove the difference.


```{r}
head(students.number.based.on.county)

```

```{r}
final_map <- left_join(istdata, students.number.based.on.county, by = "id")
head(final_map)

```


```{r}

ggplot(final_map) +geom_polygon( aes(x = long, y = lat, group = group, fill = NUMBER_OF_STUDENTS), color = "grey") +
coord_map() +theme_void() + labs(title = "Istanbul Metropolitan Municipality Distribution of scholarship winners by districts") +
scale_fill_gradient(low = "pink", high = "red",na.value = "white") +
theme(plot.title = element_text(hjust = 0.5),plot.subtitle = element_text(hjust = 0.5))

```
<br>
<br>
<br>
The plot above shows the distribution of scholarship winners by districts. The provinces are Esenyurt, Pendik and  Bağcılar has the highest number of scholarships.

```{r}
mean(ibb_yardim$ASSISTANCE_TERM)
```
The average scholarship period for scholarship winners is 1.28 term.

```{r}
mean(ibb_yardim$AGE)
```
The average age of scholarship winners is 21.35 years.

```{r}
median(ibb_yardim$AGE)
```
The median of scholarship winners' age is 21 years.

```{r}
Mode(ibb_yardim$AGE) #with capital M
```
21 is the most recurring age among scholarship recipient students.

<h1> Measure of Dispersion </h1>

<b>Range </b>


```{r}
 range<-range(ibb_yardim$AGE)
range[2]-range[1] #Shows the difference between minimum and maximum values
```
The difference between minimum age and maximum age is 10.


<b>Calculating Quantiles </b>

```{r}

quantile(ibb_yardim$AGE,0.75) #3rd quantile
```

The 75% of the students' age is 23 or below.


<b>Interquartile Range </b>

```{r}

IQR(ibb_yardim$AGE)
```
It refers to spread of 50% of the data.


<b>Variance and Standard Deviation</b>

The variance is a measure of how far each value in the data set is from the mean.Standard deviation is the measure of spread most commonly used in statistical practice when the mean is used to calculate central tendency. Thus, it measures spread around the mean.

```{r}
sd(ibb_yardim$AGE)
```
Standard Deviation is 1.82, far each age in the data set is from the average age.

<b>Covariance </b>

```{r}
cov(ibb_yardim$AGE,ibb_yardim$ASSISTANCE_TERM)

```
There are positive relationship between age and assistance time.



```{r}
chisq.test(ibb_yardim$COUNTY,ibb_yardim$UNIVERSITY)

```
Since p value is less than 0.05, It can be said that county and university are dependent variables.


<b>Measure of Skewness </b>

```{r}
library(moments) 
skewness(ibb_yardim$AGE) 
```

```{r}
hist(ibb_yardim$AGE)
```
<br>
<br>
<br>
<br>
It looks like data has bell shape.


```{r}
library(dplyr)
ages <- data.frame(ibb_yardim$AGE ) %>% group_by(ibb_yardim$AGE) %>% count(ibb_yardim$AGE) 

barplot(ages$n)
ageInt <- c(17,18,19,20,21,22,23,24,25,26,27)
axis(side = 1, at = 1:11, labels = ageInt)


```
<br>
<br>
<br>
<br>
<br>
In the figure above, the distribution of scholarship recipients by age is given. According to age distribution, the number of students at the age of 21 receiving scholarships is high.

```{r}
boxplot(ibb_yardim$AGE)
```


```{r}
counts <- table(ibb_yardim$ASSISTANCE_YEAR)
barplot(counts, main="Number of Students Distribution based on year", 
   xlab="Year")
```
<br>
<br>
<br>

According to the figure above, it can be said that more students received scholarships in 2019.

```{r}
year2019 <- ibb_yardim%>%filter(ASSISTANCE_YEAR==2019)
```


```{r}
mean(year2019$AGE)

```
The average age of the students who received a scholarship in 2019 is 21.63.





```{r}

totalStudents_2019 <-  aggregate(NUMBER_OF_STUDENTS~UNIVERSITY,year2019,sum)

head(totalStudents_2019%>%as_tibble)


```

Distribution of scholarship recipients in 2019 by university.



```{r}


ggplot(totalStudents_2019, aes(x=UNIVERSITY, y=NUMBER_OF_STUDENTS, fill=UNIVERSITY))+
  geom_bar(stat = "identity") + 
  labs(title = "Students class BarChart", x="Class", y="Response")
```




<h1> Confidence Intervals </h1>



<h3>Finding confidence interval for number of students evaluating based on year: </h3>
```{r}
year.2019<- subset(ibb_yardim, ASSISTANCE_YEAR== 2019)
year.2020 <- subset(ibb_yardim, ASSISTANCE_YEAR == 2020)
```





```{r}
Conf.2019 <-t.test(year.2019$NUMBER_OF_STUDENTS)$conf.int
Conf.2019


```

```{r}
Conf.2020 <- t.test(year.2020$NUMBER_OF_STUDENTS)$conf.int
Conf.2020

```


<h3>Finding confidence interval for age evaluating based on university:</h3>

```{r}
gtu <- subset(year2019, UNIVERSITY=="Gebze Teknik Üniversitesi")
msku <- subset(year2019, UNIVERSITY =="Muğla Sıtkı Koçman Üniversitesi")
```


```{r}
# Calculating confidence interval for age of gtü students 
gtuConf <- t.test(gtu$AGE)$conf.int
gtuConf
```


```{r}
 # Calculating confidence interval for age of mskü students 
mskuConf <- t.test(msku$AGE)$conf.int
mskuConf


```


```{r}
ibb_yardim.UNIVERSITY <- data.frame(UNIVERSITY="Gebze Teknik Üniversitesi",
                                scoreType="AGE", 
                                mean=mean(gtu$AGE), 
                                lower=gtuConf[1], 
                                upper=gtuConf[2])

```


```{r}

ibb_yardim.UNIVERSITY <- rbind(ibb_yardim.UNIVERSITY, data.frame(UNIVERSITY="Muğla Sıtkı Koçman Üniversitesi",
                                scoreType="AGE", 
                                mean=mean(msku$AGE),
                                lower=mskuConf[1], 
                                upper=mskuConf[2]))
                                



```


```{r}
ibb_yardim.UNIVERSITY
```
Summary for Muğla Sıtkı Koçman Üniversitesi and Gebze Teknik Üniversitesi.


<H1> Hypothesis Testing </H1>

<h3> Test that the true average of  MSKU students age is equal to 18. </h3>

H0: mu = 18 
Ha: mu != 18

Assesing the normality.

```{r}

shapiro.test(msku$AGE)
```
As the p-value > 0.05, the null hypothesis is not rejected, so data is normally distributed.





```{r}
t.test(msku$AGE, 
       alternative = "two.sided", 
       mu=18, 
       conf.level = 0.95)

```

To conclude, as p-value < 0.05, the null hypothesis is rejected and the confidence interval is between 21.29 and 21.81 with 95% confident. It means the average of the students age is not equal to 18.

<br>
<h3> Are	there	significant	difference	between	mean	of MSKU and GTU students' age?</h3>

Check the variance:
```{r}

x <- subset(ibb_yardim, UNIVERSITY=="Muğla Sıtkı Koçman Üniversitesi")$AGE
y <- subset(ibb_yardim, UNIVERSITY=="Gebze Teknik Üniversitesi")$AGE

var.test(x, y)



```
As p-value > 0.05, the null hypothesis is not rejected and the variances are equal.
`

```{r}
qqnorm(msku$AGE, ylab = "MSKU Students' age")
qqline(msku$AGE)
```


```{r}
t.test(x=x, y=y, alternative = "two.sided", var.equal = TRUE,
       conf.level = 0.95)
```

As p-value < 0.05, the null hypothesis is rejected and the variances are not  equal.  
We can conclude that there is statistically significant difference between msku and gtu students age.


<h3> Are	there	significant	difference	between	number of students getting scholarship based on year? </h3>


```{r}
assistance.year.2019 <- subset(ibb_yardim, ASSISTANCE_YEAR==2019)$NUMBER_OF_STUDENTS
assistance.year.2020 <- subset(ibb_yardim, ASSISTANCE_YEAR==2020)$NUMBER_OF_STUDENTS

var.test(assistance.year.2019, assistance.year.2020)


```
p-value < 0.05, the null hypothesis is  rejected and the variances are not equal.  


<h1>ANOVA </h1>
 <h3>Are	there	significant	differences	between	mean of   students' number who got scholarship  based on different university? </h3>
The Hypothesis is :   
H0: mu1=mu2=mu3=mu4=mu5=...
Ha: At least one of them is different  

```{r}
head(ibb_yardim)
```


```{r}
data <- ibb_yardim[, c(4, 7)]
names(data) <- c("UNIVERSITY", "numberOfStudents")
data$UNIVERSITY <- factor(data$UNIVERSITY)

```


```{r}
head(levels(data$UNIVERSITY))

```


```{r}
head(group_by(data, UNIVERSITY) %>%
  summarise(
    count = n(),
    mean = mean(numberOfStudents, na.rm = TRUE),
    sd = sd(numberOfStudents, na.rm = TRUE)
  ))
```





```{r}
anova_one_way <- aov(numberOfStudents~UNIVERSITY, data = data)
summary(anova_one_way)
```
The p-value is less than the usual threshold of 0.05. We can to say there is  statistical difference between the universities.





<h1> c. Results, d. Discussion </h1>

* Esenyurt, Pendik and  Bağcılar provinces has the highest number of students needs scholarships.

* It can be said that the students who receive the most scholarships are 21 years old.

* 75% of the student's age is 23 or below.

* There is a positive relationship between age and assistance time.


* The maximum age difference between scholarship recipients is 10.

* It can be said that more students received scholarships in 2019.

* We can say there is a statistical difference between the universities and the number of students who received the scholarship. 

* The sample size was very small, if the sample size can be increased, this may enable us to reach a more efficient result.

<hr class="new3">
<hr class="new3">



<h1> 2) Izmir Metropolitan Municipality Social Aid Applications </h1>

<h1>a. Data collection </h1>
<h1>Description of Data </h1>
The dataset contains the Social Assistance application information made to the Izmir metropolitan municipality.

<body>

<h2>Dataset Information</h2>

<table>
   <tr>
    <th>Column</th>
    <th>Description</th>
  </tr>
  <tr>
    <th>YIL</th>
    <th>Application year information</th>
  </tr>
  <tr>
    <td>YARDIM_TALEBI</td>
    <td>Assistance type information.</td>

  </tr>
  <tr>
    <td>BASVURU_SEKLI</td>
    <td> Application form information.</td>
  </tr>
  
  <tr>
    <td>BASVURU_TIPI</td>
    <td>application type information.</td>
  </tr>
  <tr>
    <td>CINSIYET</td>
    <td> Gender information.</td>
  </tr>
  <tr>
    <td>BASVURU_ADEDI</td>
    <td>Total number of applications</td>
  </tr>

</table>

<b> Data Source: </b> https://acikveri.bizizmir.com/dataset/9e5a897e-3988-4cc1-9f16-de35adc97431/resource/04ad0a96-4931-4491-ba85-a700ac50a7df/download/sosyal-yardim-basvurulari.csv

<b>Licence: Izmir Metropolitan Municipality License</b>
<br>
<hr>



<h1>b. Analyses </h1>
```{r}
izmir.sosyal.yardim <- read.csv("izmirSosyalYardimBasvurulari.csv")
head(izmir.sosyal.yardim)

```
The first 6 rows from the dataframe.

```{r}
izmir.sosyal.yardim <- subset(izmir.sosyal.yardim, CINSIYET !="BELİRSİZ")

head(izmir.sosyal.yardim)

```

Removing the unknown data from the gender column.


```{r}
chisq.test(izmir.sosyal.yardim$CINSIYET,izmir.sosyal.yardim$YARDIM_TALEBI)
```

Since p value is greater than 0.05, It can be said that gender and aid request are independent variables.

```{r}
ggplot(izmir.sosyal.yardim) + geom_bar(aes(x = CINSIYET))

```

Number of applications for assistance by gender. It looks men have more application.

```{r}

counts <- table(izmir.sosyal.yardim$BASVURU_TIPI)
barplot(counts, main=" DISTRIBUTION OF ASSISTANCE ", xlab="application type")
```

```{r}
summary(izmir.sosyal.yardim$BASVURU_ADEDI)
```

```{r}
dataB <- izmir.sosyal.yardim[, c("BASVURU_TIPI")]
tail(dataB)

```

```{r}

totalB <- aggregate(BASVURU_ADEDI~BASVURU_TIPI,izmir.sosyal.yardim,sum)
head(totalB)
```

Total number of applications by application type.


```{r}

ggplot(totalB, aes(x="", y=BASVURU_ADEDI, fill=BASVURU_TIPI)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  
  theme_void() 
```

Displaying the total number of applications according to the application type on the pie chart. Citizen according to chart has more application and then old people has second more application.



```{r}
total.type.assistance <- aggregate(BASVURU_ADEDI~YARDIM_TALEBI,izmir.sosyal.yardim,sum)
head(total.type.assistance)

```

Types of the assistances.


```{r}

ggplot(total.type.assistance, aes(x="", y=BASVURU_ADEDI, fill=YARDIM_TALEBI)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  
  theme_void() 
```

According to the pie chart above, most applications were made for milk aid in Izmir, followed by food aid.

```{r}
data2 <- total.type.assistance[total.type.assistance$BASVURU_ADEDI %in% tail(sort(total.type.assistance$BASVURU_ADEDI),5),]
pie(data2$BASVURU_ADEDI, data2$YARDIM_TALEBI)
```
<br>
<br>
<br>
In this pie char we can see clearly  the 5 top assistance type.

```{r}

milk_assistance<- izmir.sosyal.yardim%>%filter(YARDIM_TALEBI=="SÜT")
head(milk_assistance)


```


```{r}
total.type.assistance <- aggregate(BASVURU_ADEDI~CINSIYET,milk_assistance,sum)
total.type.assistance


```

According to the table, most people applying for milk aid are female.



```{r}
ggplot(total.type.assistance, aes(x="", y=BASVURU_ADEDI, fill=CINSIYET)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  
  theme_void() 
```


Women are the ones who demand more milk aid.


<h1> Confidence Intervals </h1>

<h3> Confidence interval for  number of applications based on gender. </h2>

```{r}
males <- subset(izmir.sosyal.yardim, CINSIYET=="ERKEK")
females <- subset(izmir.sosyal.yardim, CINSIYET=="KADIN")


```

<h4>Calculating confidence interval for number of applications of male</h4>

```{r}

malesConf <- t.test(males$BASVURU_ADEDI)$conf.int
malesConf
```

The above output indicates that the true average of number of applications for males  is between 64.74 and 112.21 with 95% confident.


<h4> Calculating confidence interval for number of applications of female </h4>

```{r}
femalesConf <- t.test(females$BASVURU_ADEDI)$conf.int
femalesConf
```

The above output indicates that the true average of score for females  is between 137.46 and 224.88 with 95% confident.

<h3> Find confidence interval for number of applications evaluating the differences based on type of applications.  </h3>


```{r}
depremzede <- subset(izmir.sosyal.yardim, BASVURU_TIPI=="DEPREMZEDE")
yaslilar <- subset(izmir.sosyal.yardim,BASVURU_TIPI=="YAŞLILAR")

```

```{r}
Confdepremzede<- t.test(depremzede$BASVURU_ADEDI)$conf.int
Confdepremzede
```

The above output indicates that the true average of number of applications for earthquake victims  is between 10.64 and 31.66 with 95% confident.


```{r}
Conf.yaslilar <- t.test(yaslilar$BASVURU_ADEDI)$conf.int
Conf.yaslilar
```

The above output indicates that the true average of number of applications for old people is between 38.84 and 100.30 with 95% confident.

<h3> Find confidence interval for number of applications evaluating the differences based on request for aid. </h3>

```{r}
food <- subset(izmir.sosyal.yardim, YARDIM_TALEBI=="GIDA")
cash.aid <- subset(izmir.sosyal.yardim, YARDIM_TALEBI=="NAKDİ YARDIM")

```

```{r}
conf.food  <- t.test(food$BASVURU_ADEDI)$conf.int
conf.food
```

The above output indicates that the true average of number of applications for those applying for food aid is between 183.20 and 391.48 with 95% confident.


```{r}
conf.cash.aid  <- t.test(cash.aid$BASVURU_ADEDI)$conf.int
conf.cash.aid
```

The above output indicates that the true average of number of applications for those applying for cash aid is between 50.16 and 197.11 with 95% confident.
<br>

<h1> Hypothesis Testing </h1>



* ASSUMPTION 1: SAMPLES ARE INDEPENDENT.
<br>
* ASSUMPTION 2: SAMPLES ARE DRWAN FROM POPULATION WITH EQUAL VARIANCES.
<br>

* ASSUMPTION 2: SAMPLES ARE DRAWN FROM POPULATIN WITH A NORMAL DISTRIBUTIONS.
<br>

<h3>Are	there	significant	difference	between	mean	of male and female people number of applications? </h3>

```{r}
females<- subset(izmir.sosyal.yardim,CINSIYET== "KADIN")
males <- subset(izmir.sosyal.yardim,CINSIYET== "ERKEK")


```


Checking for  normal distrubiton.

```{r}
shapiro.test(males$BASVURU_ADEDI)
```

As the p-value < 0.05, the null hypothesis is  rejected, so actually data is not  normally distributed.

```{r}
shapiro.test(females$BASVURU_ADEDI)
```

As the p-value < 0.05, the null hypothesis is  rejected, so actually data is not  normally distributed. But we assumed that data is normally distributed.

```{r}
t.test(izmir.sosyal.yardim$BASVURU_ADEDI ~ izmir.sosyal.yardim$CINSIYET)

```

p values is less than 0.05 so we reject to null hypothesis.It means there is significant difference between mean of females and males' number of applications.

------
<hr>
<h2>Test that, the true average of the number of aids received by celiac patients is 180. </h2>

hypothesis is formulated as following:   
H0: mu = 180 <br>

Ha: mu != 180


```{r}
colyak.patient <- subset(izmir.sosyal.yardim, BASVURU_TIPI=="ÇÖLYAK HASTASI")
head(colyak.patient)

```

```{r}
shapiro.test(colyak.patient$BASVURU_ADEDI)
```

As the p-value < 0.05, data is  not normally  distributed.

```{r}
t.test(colyak.patient$BASVURU_ADEDI, 
       alternative = "two.sided", 
       mu=180, 
       conf.level = 0.95)
```

To conclude, as p-value < 0.05, the null hypothesis is rejected.


<h2> Are	there	significant	difference	between	mean	of   male and female celiac patients  number of applications? </h2>


```{r}
qplot(x = CINSIYET, y = BASVURU_ADEDI,
      geom = "boxplot", data = colyak.patient,
      xlab = "Gender",
      ylab = "Number Of Applications",
      fill = I("lightblue"))
```


```{r}

x <- aggregate(BASVURU_ADEDI ~ CINSIYET, 
          data = colyak.patient,
          FUN = function(x) {c(mean = mean(x),
                               se = sd(x) / sqrt(length(x)))})
x
```

```{r}
x <- subset(colyak.patient, CINSIYET=="ERKEK")$BASVURU_ADEDI
y <- subset(colyak.patient, CINSIYET=="KADIN")$BASVURU_ADEDI
var.test(x, y)
```

p values is less than 0.05 so, the null hypothesis rejected variances are different.

As the sample size is not large enough, we will assess the normality.

```{r}
shapiro.test(x)

```


```{r}
t.test(x=x, y=y, alternative = "two.sided", var.equal = TRUE,
       conf.level = 0.95)
```

p value is less than 0.05, the null hyphothesis rejected. The result is said to be statistically significant.


<h1> Annova </h1>

<h3> Are	there	significant	differences	between	mean of different gender application? </h3>


```{r}
data <- izmir.sosyal.yardim[, c(6, 7)]
names(data) <- c("gender", "numberOfApplication")
data$gender <- factor(data$gender)
```

```{r}
levels(data$gender)

```

The Hypothesis:   
H0: mu1=mu2 <br>
Ha: At least one of them is different  

```{r}
model<-aov(numberOfApplication~gender, data=data)
summary(model)

```

The p-value < 0.05, we can conclude that there is  significant differences between the gender applications.   



<h1> c. Results, d. Discussion </h1>

<li> Men applied for more assistance. </li>

<li> Citizens, the disabled  people and the elderly people are the groups that apply the most for help. </li>

<li> Most applications were made for milk aid in Izmir, followed by food aid and cash aid. </li>

<li> Women citizens are the ones who demand milk aid the most. </li>

<li>I think there are no biases exist in this dataset.</li>

<li>Sample size is very small, increasing the sample size can improve this study. </li>



<br>


<hr class="new3">
<hr class="new3">
<h1>3) Plays Performed in City Theaters </h1>
<h1>a. Data collection </h1>
<h1>Description of Data </h1>
There is data on the theater plays that have been staged since September 2017. Within this data source, there is information about the theater plays, stage information, type of play, stage location and number of spectators of the City Theatres.


<html>
<head>
<style>
table {
  font-family: arial, sans-serif;
  border-collapse: collapse;
  width: 100%;
}

td, th {
  border: 1px solid #dddddd;
  text-align: left;
  padding: 8px;
}

tr:nth-child(even) {
  background-color: #dddddd;
}
</style>
</head>
<body>
<h1>a. Data collection</h1>
<h2>Dataset Information</h2>
<table>
  <tr>
    <th>Column</th>
    <th>Description</th>
  </tr>
  <tr>
    <td>PLAY_DATE</td>
    <td>t is the field containing the date that the activity will take place.  Data format is in YYYY-MM-DD format.</td>

  </tr>
  <tr>
    <td>PLAY_NAME</td>
    <td> It is the area where the names of the activities such as plays and musicals are displayed.</td>
  </tr>
  
  <tr>
    <td>THEATER_NAME</td>
    <td>It is the area where the names of the activities such as plays and musicals are displayed.</td>
  </tr>
  <tr>
    <td>PLAY_CATEGORY</td>
    <td> It contains the information in which category the game will be staged.</td>
  </tr>
  <tr>
    <td>IS_GUEST_PLAY</td>
    <td>The theater group that played the staged play contains the information whether it is from the City Theaters or the Guest. * Guest = 1 If not Guest = 0.</td>
  </tr>
  
  <tr>
    <td>LONGITUDE</td>
    <td>It contains the longitude information of the stage.</td>
  </tr>
   <tr>
    <td>LATITUDE</td>
    <td> It contains the latitude information of the stage.</td>

  </tr>
   <tr>
    <td>NUMBER_OF_AUDIENCE</td>
    <td>Indicates the number of spectators who came to the play staged.</td>

  </tr>
</table>
<b> Data Source: </b> https://data.ibb.gov.tr/dataset/cd121983-4063-4d12-a900-3c928d59d963/resource/79465ce9-8755-4b57-8e6c-def0c0caadc8/download/theater_play.csv

<b>License: Istanbul Metropolitan Municipality Open Data License </b>
<br>
<hr>


<h1>b. Analyses </h1>
```{r}
sehir.tiyatrolari <- read.csv("sehirTiyatrolari.csv")
head(sehir.tiyatrolari)

```

```{r}

sehir.tiyatrolari$PLAY_DATE <- as.numeric(substring(sehir.tiyatrolari$PLAY_DATE,0,4))

```


```{r}

year2017 <- sehir.tiyatrolari%>%filter(PLAY_DATE==2017)
year2018 <- sehir.tiyatrolari%>%filter(PLAY_DATE==2018)
year2019 <- sehir.tiyatrolari%>%filter(PLAY_DATE==2019)
year2020 <- sehir.tiyatrolari%>%filter(PLAY_DATE==2020)


mean(year2017$NUMBER_OF_AUDIENCE)
mean(year2018$NUMBER_OF_AUDIENCE)
mean(year2019$NUMBER_OF_AUDIENCE)
mean(year2020$NUMBER_OF_AUDIENCE)



```

```{r}
summary(sehir.tiyatrolari$NUMBER_OF_AUDIENCE)
```

```{r}
counts <- table(sehir.tiyatrolari$PLAY_DATE)

head(counts)
```

```{r}
barplot(counts, main="  Audience and Years")
```


```{r}


ggplot(sehir.tiyatrolari,aes(x=	
PLAY_CATEGORY,y=NUMBER_OF_AUDIENCE))+geom_point()+labs(title = "The relationship between number of audience  and play category")

```

```{r}
play.based.on.category <- aggregate(NUMBER_OF_AUDIENCE~PLAY_CATEGORY,sehir.tiyatrolari,sum)

ggplot(play.based.on.category, aes(x="", y=NUMBER_OF_AUDIENCE, fill=PLAY_CATEGORY)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  
  theme_void()

```

Adult plays have a maximum audience.

```{r}

play.type <- aggregate(NUMBER_OF_AUDIENCE~PLAY_TYPE,sehir.tiyatrolari,sum)

ggplot(play.type, aes(x="", y=NUMBER_OF_AUDIENCE, fill=PLAY_TYPE)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  
  theme_void() 

```

<h1> Confidence Intervals </h1>


<h3>Finding confidence interval for number of audience  evaluating the differences based on play type.</h3>


```{r}
localPlay <- subset(sehir.tiyatrolari, PLAY_TYPE=="Yerli")
foreginPlay <- subset(sehir.tiyatrolari, PLAY_TYPE=="Yabancı")
```

```{r}
# Calculating confidence interval for number of audience of local plays
localConf <- t.test(localPlay$NUMBER_OF_AUDIENCE)$conf.int
localConf
```

The above output indicates that the true average of number of audiences for  local play is between 60.01 and 64.69 with 95% confident.

```{r}
 foreignConf <- t.test(foreginPlay$NUMBER_OF_AUDIENCE)$conf.int
 foreignConf
```

The above output indicates that the true average of number of audience for foreign plays' is between 68.04 and 74.88 with 95% confident.


```{r}
local.play.audience <- data.frame( PLAY_TYPE="Yerli",
                                mean=mean(localPlay$NUMBER_OF_AUDIENCE), 
                                lower=localPlay[1], 
                                upper=localPlay[2])

play.audience<- rbind(local.play.audience, data.frame(PLAY_TYPE="Yabancı",
                                mean=mean(foreginPlay$NUMBER_OF_AUDIENCE), 
                                lower=foreginPlay[1], 
                                upper=foreginPlay[2]))


ggplot(play.audience, aes("NUMBER_OF_AUDIENCE", mean)) + geom_point()

```

<h3>Finding confidence interval for number of audience evaluating the differences based on play date. </h3>

```{r}
year2017 <- subset(sehir.tiyatrolari, PLAY_DATE==2017)
year2020 <- subset(sehir.tiyatrolari, PLAY_DATE==2020)

```

```{r}
conf2017 <- t.test(year2017$NUMBER_OF_AUDIENCE)$conf.int
conf2017
```

The above output indicates that the true average of number of audince for 2017 year is between 78.38 and 89.53 with 95% confident

```{r}
year2020 <- t.test(year2020$NUMBER_OF_AUDIENCE)$conf.int
year2020
```

The above output indicates that the true average of number of audince for 2020 year is between 730.13 and 89.53 with 95% confident.

<h3>Finding confidence interval for number of audience evaluating the differences based on theatre. </h3>

```{r}
kadikoyTheatre <-  subset(sehir.tiyatrolari, THEATER_NAME=="Kadıköy Haldun Taner Sahnesi")
umraniyeTheatre<- subset(sehir.tiyatrolari, THEATER_NAME=="Ümraniye Sahnesi")

```

```{r}
confKadikoy <- t.test(kadikoyTheatre$NUMBER_OF_AUDIENCE)$conf.int
confKadikoy
```


The above output indicates that the true average of number of audience for "Kadıköy Haldun Taner Sahnesi"  is between 69.33 and 88. with 95% confident.
```{r}
confUmraniye <- t.test(umraniyeTheatre$NUMBER_OF_AUDIENCE)$conf.int
confUmraniye
```
The above output indicates that the true average of number of audince for "Ümraniye Sahnesi"  is between 66.69 and 78.23 with 95% confident.


```{r}
head(sehir.tiyatrolari)
```

<h1>HYPHOTESIS TESTING <h1>

<h3> Let's test that the true average number of audience of local plays mean is  900. </h3>

H0: mu = 900   
Ha: mu != 900




```{r}
t.test(localPlay$NUMBER_OF_AUDIENCE, 
       alternative = "two.sided", 
       mu=900, 
       conf.level = 0.95)
```

<p>To conclude, as p-value < 0.05, the null hypothesis is  rejected and the confidence interval is not  between 60.01 and 64.69 with 95% confident. </p>


<h3> Are	there	significant	difference	between	mean	of Turkish	and	foreigner play category based on audience number? </h3>

```{r}
yerli <- subset(sehir.tiyatrolari, PLAY_TYPE=="Yerli")$NUMBER_OF_AUDIENCE
yabanci <- subset(sehir.tiyatrolari, PLAY_TYPE=="Yabancı")$NUMBER_OF_AUDIENCE
var.test(yerli, yabanci)
```

<p>As p-value < 0.05, the null hypothesis is  rejected and the variances are not equal.  

As the sample size is not large enough, we will assess the normality. </p>

```{r}
yerliOyun <- subset(sehir.tiyatrolari,PLAY_TYPE=="Yerli")$NUMBER_OF_AUDIENCE
x <- t.test(yerliOyun)$conf.int
x

```

<p>The above output indicates that the true average of number of audience for local play is between 60 and 64 with 95% confident. <p>

```{r}
yabanciOyun  <- subset(sehir.tiyatrolari,PLAY_TYPE=="Yabancı")$NUMBER_OF_AUDIENCE
y <- t.test(yabanciOyun)$conf.int
y
```

<p>The above output indicates that the true average of number of audience for foreign play is  between  68 and 74 with 95% confident. </p>


<br>

<h3> Test that the true average number of audience of Musical play mean is  150. </h3>
H0: mu = 150   
Ha: mu != 150

```{r}


musical.play <- subset(sehir.tiyatrolari,PLAY_CATEGORY=="Müzikal")


```

```{r}
shapiro.test(musical.play$NUMBER_OF_AUDIENCE)

```
As the p-value < 0.05, so data is not normally distributed. But we assumed that the data is normally distributed.

```{r}
qqnorm(musical.play$NUMBER_OF_AUDIENCE)
qqline(musical.play$NUMBER_OF_AUDIENCE)

```

```{r}
t.test(musical.play$NUMBER_OF_AUDIENCE, 
       alternative = "two.sided", 
       mu=150, 
       conf.level = 0.95)
```

To conclude, as p-value < 0.05, the null hypothesis is rejected. 

<h1>ANOVA </h1>
<h3> Are	there	significant	differences	between	mean of different categories of audiences? </h3>

```{r}
head(sehir.tiyatrolari)
```

```{r}
data <- sehir.tiyatrolari[, c(5, 10)]
names(data) <- c("category", "numberOfAudience")
data$category <- factor(data$category)
head(data)
```


```{r}
levels(data$category)

```

```{r}
group_by(data, category) %>%
  summarise(
    count = n(),
    mean = mean(numberOfAudience, na.rm = TRUE),
    sd = sd(numberOfAudience, na.rm = TRUE)
  )
```

The Hypothesis is as following:   
H0: mu1=mu2=mu3
Ha: At least one of them is different   


```{r}
ggplot(data, aes(x = category, y = numberOfAudience, fill = category)) +
    geom_boxplot() +
    geom_jitter(shape = 15,
        color = "red",
        position = position_jitter(0.21)) +
    theme_classic()
```


```{r}
anova_one_way <- aov(numberOfAudience~category, data = data)
summary(anova_one_way)
```

The p-value is less than the usual threshold of 0.05. We can to say there is statistical difference between the categories.


<h1> c. Results, d. Discussion </h1>

<li>According to the data set, it is seen that the number of viewers has decreased considerably in 2020. </li>

<li> Istanbul city theaters had the highest audience in 2019. </li>

<li>Among the games exhibited, the most interesting adult category was shown. </li>

<li>In Istanbul city theaters, the audience showed the most interest in the local play type. </li>

<li>Sample size is very small, increasing the sample size can improve this study.</li>

<li>I think there are no biases exist in this dataset.</li>




<br>
<br>
<br>

