---
title: "izmir-sosyalYardimBasvurulari"
output: html_document
---

<h1>Description of Data </h1>
The dataset contains the Social Assistance application information made to the Izmir metropolitan municipality.


<html>
<head>
<style>
table {
  font-family: arial, sans-serif;
  border-collapse: collapse;
  width: 100%;
}

td, th {
  border: 1px solid #dddddd;
  text-align: left;
  padding: 8px;
}

tr:nth-child(even) {
  background-color: #dddddd;
}
</style>
</head>
<body>

<h2>Dataset Information</h2>

<table>
   <tr>
    <th>Column</th>
    <th>Description</th>
  </tr>
  <tr>
    <th>YIL</th>
    <th>Application year information</th>
  </tr>
  <tr>
    <td>YARDIM_TALEBI</td>
    <td>Assistance type information.</td>

  </tr>
  <tr>
    <td>BASVURU_SEKLI</td>
    <td> Application form information.</td>
  </tr>
  
  <tr>
    <td>BASVURU_TIPI</td>
    <td>application type information.</td>
  </tr>
  <tr>
    <td>CINSIYET</td>
    <td> Gender information.</td>
  </tr>
  <tr>
    <td>BASVURU_ADEDI</td>
    <td>Total number of applications</td>
  </tr>

</table>

<b>Licence: Izmir Metropolitan Municipality License</b>
<br>
<hr>

```{r}
library(ggplot2)
library(dplyr) 
```




```{r}
izmir.sosyal.yardim <- read.csv("izmirSosyalYardimBasvurulari.csv")
head(izmir.sosyal.yardim)

```

The first 6 rows from the dataframe.

```{r}
izmir.sosyal.yardim <- subset(izmir.sosyal.yardim, CINSIYET !="BELİRSİZ")

head(izmir.sosyal.yardim)

```
Removing the unknown data from the gender column.
```{r}
chisq.test(izmir.sosyal.yardim$CINSIYET,izmir.sosyal.yardim$YARDIM_TALEBI)
```
Since p value is greater than 0.05, It can be said that gender and aid request are independent variables.

```{r}
ggplot(izmir.sosyal.yardim) + geom_bar(aes(x = CINSIYET))

```
Number of applications for assistance by gender. It looks men have more application.

```{r}

counts <- table(izmir.sosyal.yardim$BASVURU_TIPI)
barplot(counts, main=" DISTRIBUTION OF ASSISTANCE ", xlab="application type")
```


```{r}
summary(izmir.sosyal.yardim$BASVURU_ADEDI)
```


```{r}
dataB <- izmir.sosyal.yardim[, c("BASVURU_TIPI")]
tail(dataB)

```

```{r}

totalB <- aggregate(BASVURU_ADEDI~BASVURU_TIPI,izmir.sosyal.yardim,sum)
head(totalB)
```
Total number of applications by application type.

```{r}

ggplot(totalB, aes(x="", y=BASVURU_ADEDI, fill=BASVURU_TIPI)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  
  theme_void() 
```
Displaying the total number of applications according to the application type on the pie chart. Citizen according to chart has more application and then old people has second more application.


```{r}
total.type.assistance <- aggregate(BASVURU_ADEDI~YARDIM_TALEBI,izmir.sosyal.yardim,sum)
head(total.type.assistance)

```
Types of the assistances.

```{r}
# Basic piechart
ggplot(total.type.assistance, aes(x="", y=BASVURU_ADEDI, fill=YARDIM_TALEBI)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  
  theme_void() # remove background, grid, numeric labels
```
According to the pie chart above, most applications were made for milk aid in Izmir, followed by food aid.

```{r}
data2 <- total.type.assistance[total.type.assistance$BASVURU_ADEDI %in% tail(sort(total.type.assistance$BASVURU_ADEDI),5),]
pie(data2$BASVURU_ADEDI, data2$YARDIM_TALEBI)
```
In this pie char we can see clearly  the 5 top assistance type.

```{r}

milk_assistance<- izmir.sosyal.yardim%>%filter(YARDIM_TALEBI=="SÜT")
head(milk_assistance)


```


```{r}
total.type.assistance <- aggregate(BASVURU_ADEDI~CINSIYET,milk_assistance,sum)
total.type.assistance


```
According to the table, most people applying for milk aid are female.

```{r}
# Basic piechart
ggplot(total.type.assistance, aes(x="", y=BASVURU_ADEDI, fill=CINSIYET)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  
  theme_void() # remove background, grid, numeric labels
```

Women are the ones who demand more milk aid.





<h1> Confidence Intervals </h1>

<h2> 1)confidence interval for students number of applications based on gender. </h2>

```{r}
males <- subset(izmir.sosyal.yardim, CINSIYET=="ERKEK")
females <- subset(izmir.sosyal.yardim, CINSIYET=="KADIN")


```
<h4>Calculating confidence interval for number of applications of male   </h4>

```{r}

malesConf <- t.test(males$BASVURU_ADEDI)$conf.int
malesConf
```
The above output indicates that the true average of number of applications for males  is between 64.74 and 112.21 with 95% confident.


<h4> Calculating confidence interval for number of applications of female <h4>
```{r}
femalesConf <- t.test(females$BASVURU_ADEDI)$conf.int
femalesConf
```
The above output indicates that the true average of score for females  is between 137.46 and 224.88 with 95% confident.





<h2>2) Find confidence interval for number of applications evaluating the differences based on type of applications.  </h2>

```{r}
depremzede <- subset(izmir.sosyal.yardim, BASVURU_TIPI=="DEPREMZEDE")
yaslilar <- subset(izmir.sosyal.yardim,BASVURU_TIPI=="YAŞLILAR")

```



```{r}
Confdepremzede<- t.test(depremzede$BASVURU_ADEDI)$conf.int
Confdepremzede
```

The above output indicates that the true average of number of applications for earthquake victims  is between 10.64 and 31.66 with 95% confident.

```{r}
Conf.yaslilar <- t.test(yaslilar$BASVURU_ADEDI)$conf.int
Conf.yaslilar
```
The above output indicates that the true average of number of applications for old people is between 38.84 and 100.30 with 95% confident.



<h2> 3) Find confidence interval for number of applications evaluating the differences based on request for aid. </h2>

```{r}
food <- subset(izmir.sosyal.yardim, YARDIM_TALEBI=="GIDA")
cash.aid <- subset(izmir.sosyal.yardim, YARDIM_TALEBI=="NAKDİ YARDIM")

```



```{r}
conf.food  <- t.test(food$BASVURU_ADEDI)$conf.int
conf.food
```
The above output indicates that the true average of number of applications for those applying for food aid is between 183.20 and 391.48 with 95% confident.

```{r}
conf.cash.aid  <- t.test(cash.aid$BASVURU_ADEDI)$conf.int
conf.cash.aid
```
The above output indicates that the true average of number of applications for those applying for cash aid is between 50.16 and 197.11 with 95% confident.




<h1> Hyphotesis Testing </h1>



ASSUMPTION 1: SAMPLES ARE INDEPENDENT.
ASSUMPTION 2: SAMPLES ARE DRWAN FROM POPULATION WITH EQUAL VARIANCES.
ASSUMPTION 2: SAMPLES ARE DRAWN FROM POPULATIN WITH A NORMAL DISTRIBUTIONS.

<h2>Are	there	significant	difference	between	mean	of male and female people number of applications? <h2>


```{r}
females<- subset(izmir.sosyal.yardim,CINSIYET== "KADIN")
males <- subset(izmir.sosyal.yardim,CINSIYET== "ERKEK")


```


check normal distrubiton

```{r}
shapiro.test(males$BASVURU_ADEDI)
```
As the p-value < 0.05, the null hypothesis is  rejected, so actually data is not  normally distributed.

```{r}
shapiro.test(females$BASVURU_ADEDI)
```
As the p-value < 0.05, the null hypothesis is  rejected, so actually data is not  normally distributed. But we assumed that data is normally distributed.


```{r}
t.test(izmir.sosyal.yardim$BASVURU_ADEDI ~ izmir.sosyal.yardim$CINSIYET)

```
p values is less than 0.05 so we reject to null hyphotehesis.It means there is significant difference between mean of females and males' number of applications.

------
<hr>
<h2>the true average of the number of aids received by celiac patients is 180. </h2>

hypothesis is formulated as following:   
H0: mu = 180
Ha: mu != 180




```{r}
colyak.patient <- subset(izmir.sosyal.yardim, BASVURU_TIPI=="ÇÖLYAK HASTASI")
head(colyak.patient)

```


```{r}
shapiro.test(colyak.patient$BASVURU_ADEDI)
```
As the p-value < 0.05, data is  not normally  distributed.

```{r}
t.test(colyak.patient$BASVURU_ADEDI, 
       alternative = "two.sided", 
       mu=180, 
       conf.level = 0.95)
```
To conclude, as p-value < 0.05, the null hypothesis is rejected.


<h2> Are	there	significant	difference	between	mean	of   male and female celiac patients  number of applications? <h2>



```{r}
qplot(x = CINSIYET, y = BASVURU_ADEDI,
      geom = "boxplot", data = colyak.patient,
      xlab = "Gender",
      ylab = "Score",
      fill = I("lightblue"))
```

```{r}

x <- aggregate(BASVURU_ADEDI ~ CINSIYET, 
          data = colyak.patient,
          FUN = function(x) {c(mean = mean(x),
                               se = sd(x) / sqrt(length(x)))})
x
```
```{r}
x <- subset(colyak.patient, CINSIYET=="ERKEK")$BASVURU_ADEDI
y <- subset(colyak.patient, CINSIYET=="KADIN")$BASVURU_ADEDI
var.test(x, y)
```
p values is less than 0.05 so, the null hypothesis rejected variances are different.

As the sample size is not large enough, we will assess the normality.
```{r}
shapiro.test(x)

```

```{r}
t.test(x=x, y=y, alternative = "two.sided", var.equal = TRUE,
       conf.level = 0.95)
```
p value is less than 0.05, the null hyphothesis rejected. The result is said to be statistically significant.

```{r}
head(izmir.sosyal.yardim)
```
<h1> Annova </h1>

Q1: Are	there	significant	differences	between	mean of different gender application?
```{r}
data <- izmir.sosyal.yardim[, c(6, 7)]
names(data) <- c("gender", "numberOfApplication")
data$gender <- factor(data$gender)
```

```{r}
levels(data$gender)

```
The Hypothesis:   
H0: mu1=mu2
Ha: At least one of them is different  
 
```{r}
model<-aov(numberOfApplication~gender, data=data)
summary(model)

```
 

The p-value < 0.05, we can conclude that there is  significant differences between the gender applications.   


<h1> DISCUSSION: </h1>

<li> Men applied for more assistance. 

<li> Citizens, the disabled  people and the elderly people are the groups that apply the most for help. </li>

<li> Most applications were made for milk aid in Izmir, followed by food aid and cash aid. </li>

<li> Women citizens are the ones who demand milk aid the most. </li>

<li>I think there are no biases exist in this dataset.</li>

<li>Sample size is very small, increasing the sample size can improve this study. </li>








